---
title: "assignment_6_Anna_Rescifina"
author: "Anna Rescifina"
format: html
editor: visual
---

## SetUp

```{r}
#SetUp 

library(DBI)
library(duckdb)
# library(nycflights13)
library(tidyverse)
library(dbplyr)

conn <- dbConnect(duckdb(), "data/flights.duckdb", read_only = TRUE)

dbIsValid(conn)
dbListTables(conn)
dbListFields(conn, "flights")
```

## Part 1a

```{r}
# Part 1a
dbGetQuery(conn, "
SELECT *
FROM flights
LIMIT 10;
")
```

## Part 1b

```{r}
# Part 1b (i)
dbGetQuery(conn, "
SELECT COUNT(*) AS total_flights
FROM flights;
")
```

This query calculates the total number of flights in the dataset using COUNT(\*), which counts all rows in the flights table. The result is stored in a new column called total_flights, providing a quick overview of the dataset size

## Part 1b

```{r}
# Part 1b (ii)
dbGetQuery(conn, "
SELECT carrier, COUNT(*) AS total_flights_by_carrier
FROM flights
GROUP BY carrier
ORDER BY total_flights_by_carrier DESC;
")
```

This query calculates the total number of flights for each carrier. By using GROUP BY carrier, the data is grouped by airline, and COUNT(\*) counts how many flights each carrier operated. The results are sorted in descending order using ORDER BY allowing us to easily identify which carriers have the highest number of flights

## Part 1c

```{r}
# Part 1c
dbGetQuery(conn, "
SELECT AVG(dep_delay) AS avg_dep_delay
FROM flights;
")
```

This query computes the average departure delay across all flights using the AVG(dep_delay) function. The result is returned as a single value in a column named avg_dep_delay , which summarizes overall departure performance in the dataset

## Part 1d

```{r}
# Part 1d
dbGetQuery(conn, "
SELECT dest, COUNT(*) AS total_flights
FROM flights
GROUP BY dest
ORDER BY total_flights DESC
LIMIT 5;
")
```

This query identifies the top 5 most frequent destinations in the dataset. It groups the data by destination using GROUP BY best, counts the number of flights to each destination, and sorts the results in descending order. The LIMIT 5 statement ensures that only the top five destinations with the highest number of flights are displayed.

## Part 1e

```{r}
# Part 1e
avg_delay_by_carrier <- dbGetQuery(conn, "
SELECT carrier, AVG(arr_delay) AS avg_arr_delay
FROM flights
GROUP BY carrier
ORDER BY avg_arr_delay DESC;
")

avg_delay_by_carrier
```

This query calculates the average arrival delay for each carrier. Using GROUP BY carrier , the data is grouped by airline, and AVG(arr_delay) computes the average delay for each group. The results are sorted in descending order to highlight which carriers experience the highest average arrival delays. The output is stored as a table for further analysis

## Disconnect

```{r}
dbDisconnect(conn)
dbIsValid(conn)
```

```{r}
dbDisconnect(conn, shutdown = TRUE)
```
